quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-11-11 paper-quollr.Rmd
Will change after comments later

The `gen_centroids()` function calculates the centroids of a hexagonal grid.

The coordinate limits of the embedding (`lim1` and `lim2`) are used to compute the aspect ratio between the two axes, which informs vertical spacing. The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagons along y-axis (`b2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`b1`) and buffer (`q`).

Then, the centroids are computed iteratively. The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width. Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect. Vertical spacing (`vs`) is given by $\sqrt{3}/2 \times a_1$.

The y-coordinates for each row are similarly calculated, and paired with the x-coordinates based on whether the total number of rows is even or odd. In the case of an odd number of rows, the final row uses only the odd-row x-coordinates to maintain the alternating pattern.

Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2\text{-}D$ space.

@@

Each hexagon is defined relative to its centroid $(C_x, C_y)$, with six vertices positioned equidistantly around the center. The function first verifies the presence of the required hexagon width parameter `a1`. This width determines the horizontal spacing (`hs`).

Two derived constants are calculated to define the relative distances to the vertices. The horizontal and vertical offset is defined as $dx = a_1/2$, and $dy = a_1/\sqrt{3}$ repectively. A vertical spacing factor $vf = a_1/2\sqrt{3}$ refines vertical placement in staggered rows.

With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid. These offsets form two vectors corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.

For each centroid, six vertices are computed and assigned a polygon ID as the centroid. These vertices are then combined into a tibble that records the polygon ID (`h`) and the respective x (`x`) and y (`y`) coordinates for all hexagon corners.

To reduce computational overhead, the geometry calculations are implemented in C++ using `gen_hex_coord_cpp()`, which returns a `tibble` of vertex coordinates.

>>>>

The `gen_centroids()` function computes the centroids of a hexagonal grid. The function first validates the input parameters, ensuring that the number of bins along the x-axis (`b1`) is at least two and that the buffer width (`q`) lies within the recommended range of 0.05–0.2.

The coordinate limits of the embedding (`lim1` and `lim2`) are used to calculate the aspect ratio between the two axes, which determines the vertical scaling of the grid. The helper function `calc_bins_y()` is then called to compute the number of bins along the y-axis (`b2`) and the hexagon binwidth (`a1`) given `b1` and `q`. The hexagon grid starting points are defined by `s1` and `s2`. These are set to $-q$ and $-q \times r2$, where `r2` is the is the aspect ratio between the two axes.

Centroids are generated by alternating the x-coordinates of adjacent rows to create the hexagonal tiling. Odd-numbered rows use x-coordinates spaced by the hexagon width (`a1`), while even-numbered rows are staggered by half this distance (`a1/2`). The vertical spacing between rows is $v_s = (\sqrt{3}/2) a_1$.

The y-coordinates are constructed using this spacing, and paired with x-coordinates depending on whether the total number of rows (`b2`) is even or odd. For an odd number of rows, the final row includes only the odd-row x-coordinates to preserve the alternating pattern.

The output is a tibble containing a unique hexagon ID (`h`) and centroid coordinates (`c_x`, `c_y`), which together define the layout of the complete hexagonal grid over the $2\text{-}D$ space.

@@

The function first verifies that the hexagon width (`a1`) is provided, as it determines the overall scale of the grid. Each hexagon is defined relative to its centroid $(C_x, C_y)$, with six vertices positioned symmetrically around the center.

Three constants are derived to define the relative distances to the vertices:
$dx = a_1/2, \quad dy = a_1/\sqrt{3}, \quad vf = a_1/(2\sqrt{3})$.

Here, `dx` and `dy` define the horizontal and vertical offsets, while `vf` refines vertical positioning to maintain geometric alignment across rows.

Using these constants, two offset vectors are defined for the six vertices relative to the centroid: $x_{\text{add}} = (0, -dx, -dx, 0, dx, dx), \quad y_{\text{add}} = (dy, vf, -vf, -dy, -vf, vf)$.

These offsets correspond to the six directions around each centroid: top, top-left, bottom-left, bottom, bottom-right, and top-right.

For each centroid, the coordinates of its six vertices are computed as $x_i = C_x + x_{\text{add},i}, \quad y_i = C_y + y_{\text{add},i}$, and assigned a polygon ID equal to the centroid identifier (`h`). The resulting coordinates are combined into a tibble containing the polygon ID (`h`) and vertex coordinates (`x`, `y`).

To improve computational efficiency, the vertex generation is implemented in C++ via the helper function `gen_hex_coord_cpp()`. This routine iterates over all centroids, applies the vertex offsets, and returns a tibble that defines the complete set of hexagon geometries in $2\text{-}D$ space.

### 2025-11-11 paper-quollr.tex
Will change after comments later

The \texttt{gen\_centroids()} function calculates the centroids of a hexagonal grid.

The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing. The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagons along y-axis (\texttt{b2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{b1}) and buffer (\texttt{q}).

Then, the centroids are computed iteratively. The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width. Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect. Vertical spacing (\texttt{vs}) is given by \(\sqrt{3}/2 \times a_1\).

The y-coordinates for each row are similarly calculated, and paired with the x-coordinates based on whether the total number of rows is even or odd. In the case of an odd number of rows, the final row uses only the odd-row x-coordinates to maintain the alternating pattern.

Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2\text{-}D\) space.

@@

Each hexagon is defined relative to its centroid \((C_x, C_y)\), with six vertices positioned equidistantly around the center. The function first verifies the presence of the required hexagon width parameter \texttt{a1}. This width determines the horizontal spacing (\texttt{hs}).

Two derived constants are calculated to define the relative distances to the vertices. The horizontal and vertical offset is defined as \(dx = a_1/2\), and \(dy = a_1/\sqrt{3}\) repectively. A vertical spacing factor \(vf = a_1/2\sqrt{3}\) refines vertical placement in staggered rows.

With these values, the function determines fixed offsets in the x and y directions for all six vertices relative to the centroid. These offsets form two vectors corresponding to the six compass directions used to define the polygon shape: top, top-left, bottom-left, bottom, bottom-right, and top-right.

For each centroid, six vertices are computed and assigned a polygon ID as the centroid. These vertices are then combined into a tibble that records the polygon ID (\texttt{h}) and the respective x (\texttt{x}) and y (\texttt{y}) coordinates for all hexagon corners.

To reduce computational overhead, the geometry calculations are implemented in C++ using \texttt{gen\_hex\_coord\_cpp()}, which returns a \texttt{tibble} of vertex coordinates.

>>>>

The \texttt{gen\_centroids()} function computes the centroids of a hexagonal grid. The function first validates the input parameters, ensuring that the number of bins along the x-axis (\texttt{b1}) is at least two and that the buffer width (\texttt{q}) lies within the recommended range of 0.05--0.2.

The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to calculate the aspect ratio between the two axes, which determines the vertical scaling of the grid. The helper function \texttt{calc\_bins\_y()} is then called to compute the number of bins along the y-axis (\texttt{b2}) and the hexagon binwidth (\texttt{a1}) given \texttt{b1} and \texttt{q}. The hexagon grid starting points are defined by \texttt{s1} and \texttt{s2}. These are set to \(-q\) and \(-q \times r2\), where \texttt{r2} is the is the aspect ratio between the two axes.

Centroids are generated by alternating the x-coordinates of adjacent rows to create the hexagonal tiling. Odd-numbered rows use x-coordinates spaced by the hexagon width (\texttt{a1}), while even-numbered rows are staggered by half this distance (\texttt{a1/2}). The vertical spacing between rows is \(v_s = (\sqrt{3}/2) a_1\).

The y-coordinates are constructed using this spacing, and paired with x-coordinates depending on whether the total number of rows (\texttt{b2}) is even or odd. For an odd number of rows, the final row includes only the odd-row x-coordinates to preserve the alternating pattern.

The output is a tibble containing a unique hexagon ID (\texttt{h}) and centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which together define the layout of the complete hexagonal grid over the \(2\text{-}D\) space.

@@

The function first verifies that the hexagon width (\texttt{a1}) is provided, as it determines the overall scale of the grid. Each hexagon is defined relative to its centroid \((C_x, C_y)\), with six vertices positioned symmetrically around the center.

Three constants are derived to define the relative distances to the vertices:
\(dx = a_1/2, \quad dy = a_1/\sqrt{3}, \quad vf = a_1/(2\sqrt{3})\).

Here, \texttt{dx} and \texttt{dy} define the horizontal and vertical offsets, while \texttt{vf} refines vertical positioning to maintain geometric alignment across rows.

Using these constants, two offset vectors are defined for the six vertices relative to the centroid: \(x_{\text{add}} = (0, -dx, -dx, 0, dx, dx), \quad y_{\text{add}} = (dy, vf, -vf, -dy, -vf, vf)\).

These offsets correspond to the six directions around each centroid: top, top-left, bottom-left, bottom, bottom-right, and top-right.

For each centroid, the coordinates of its six vertices are computed as \(x_i = C_x + x_{\text{add},i}, \quad y_i = C_y + y_{\text{add},i}\), and assigned a polygon ID equal to the centroid identifier (\texttt{h}). The resulting coordinates are combined into a tibble containing the polygon ID (\texttt{h}) and vertex coordinates (\texttt{x}, \texttt{y}).

To improve computational efficiency, the vertex generation is implemented in C++ via the helper function \texttt{gen\_hex\_coord\_cpp()}. This routine iterates over all centroids, applies the vertex offsets, and returns a tibble that defines the complete set of hexagon geometries in \(2\text{-}D\) space.
